Index measures for oak decline severity using phenotypic descriptors

### 2019-12-06 manuscript/manuscript.Rmd
added a few notes

- define phenotype and descriptor
- Acute and chronic decline associated phenotypes
- The oak decline spectrum move away from simplistic binary categorisations
- Manual assignment of decline status difficult prone to error for due to contribution of many factors.

@@

Here, we propose a framework for quantitatively measuring the severity of oak decline

@@

- Site descriptions
- Field phenotyping protocol

Phenotypic data were collected from a total of 174 trees from 9 sites from across the south, eas and west of England.

@@

- Aggregation of cardinal observations.
- Random forest and MDS analyses

R v`r str_c(version$major,version$minor,sep ='.')` was used for all parsing, preparation and analysis of phenotypic data.

@@

- Descriptions of collected phenotypic descriptors and manual status assignment

@@

- Analysis of cross site differences
- Outline of cross site correction approach using mean centering adjustment

@@

### Phenotypic decline and decline acuteness indexes for oak decline severity

- Decline index calculation
- Comparison of indexes with manually assigned decline status labels
- Top ranked descriptor index response surfaces

Using unsupervised random forest proximities and the phenotypic descriptors as reported previously,multidimensional scaling (MDS) to two dimensions was performed to yield two phenotypic indexes:

* **Phenotypic decline index (PDI)** - the inverse of the phenotypic health index reported previously providing a measure of how severely a tree is declining. 
Min-max scaling performed on the first MDS dimension giving score between 0 and 1, with more severely declining trees having a score closer to 1.

@@

Crown related descriptors were found to contribute most to the PDI (Figure 2a) with more severely declining trees having reduced crown volume and condition. 
The DAI was more complex in descriptor contribution with trunk staining and bleeding descriptors contributing highly with bleed prevalence contributing the most. Trees with any black staining or bleeding had scores greater than 0. 
Trees with high bleed prevalence and therefore a larger area of their trunk covered with black staining or active bleeds had much higher DAI scores.

@@

### Trends of key phenotypic descriptors within the PDI and DAI measures

>>>>

- define phenotype and descriptor
- Acute and chronic decline associated phenotypes
- The oak decline spectrum move away from simplistic binary categorisations
- Manual assignment of decline status difficult prone to error for due to contribution of many factors. 
- Random forest
- Unsupervised learning for complex patterns between descriptors.

@@

Random forest is an ensemble machine learning

Unsupervised learning ...

Here, we propose a framework for quantitatively measuring the phenotypic severity of oak decline that is based on unsupervised random forest machine learning

@@

- Site descriptions
- Field phenotyping protocol
- Crown assessments
- Cardinal assessments
- Stem assessments - 2m base, 3m base

Phenotypic data were collected from a total of 174 trees from 9 sites from across the south, east and west of England (Figure #).
A total of 30 phenotypic descriptors were collected relating to the trees shape, size, crown condition and the presence of biotic agents including insects, bacteria and fungi.

@@

- Aggregation of cardinal observations.
- Random forest and MDS analyses
- Descriptions of collected phenotypic descriptors and manual status assignment

R v`r str_c(version$major,version$minor,sep ='.')` was used for all parsing, preparation and analysis of phenotypic data (ref).

Cardinal crown radius measurements were averaged.

@@

@@

@@

### Calculation of indexes for oak decline severity

- Decline index calculation
- Comparison of indexes with manually assigned decline status labels
- Top ranked descriptor index response surfaces
- Random forest able to assess categorical and continuous descriptors simultaneously

The phenotypic descriptors were analysed using 100 repeatitions of unsupervised random forest.
The proximity 
The relative proximities and the phenotypic descriptors multidimensional scaling (MDS) to two dimensions was performed to yield two phenotypic indexes:

Proximity matrix transformed to a disimilarity (1- P) then multidimensional scaling (MDS) to 2 dimensions applied.
The first dimension was found to be related to the overall severity of decline. 
It was min-max scaled to give the PDI a scale between 0 and 1 with more severly declining trees having a score closer to 1.

@@

### Descriptor contributions and trends

To understand the descriptors contributing to each of the indexes, individual random forest regression models of the phenotypic descriptors were trained, supervised using each decline index. 
Crown related descriptors were found to contribute most to the PDI (Figure #) with more severely declining trees having reduced crown volume and condition. 
The DAI was a more complex in descriptor contribution with trunk staining and bleeding descriptors contributing highly with bleed prevalence contributing the most. Trees with any black staining or bleeding had scores greater than 0. 
Trees with high bleed prevalence and therefore a larger area of their trunk covered with black staining or active bleeds had much higher DAI scores.

@@
